---
title: "R Notebook"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
#Only install these packages one time! Then either delete this chunk or comment it out!
# install.packages("devtools")
# devtools::install_github("ucl-cssb/flopr")
```

```{r load and other packages}

# Packages To Install -----------------------------------------------------
source("../Functions/Load_Packages.R")
```

```{r load thematic elements}
# thematic elements -----------------------------------------------------
source("../Functions/SKS_Personal_Theme.R")
```

## Loading Data

```{r}

#using the flopr package to wrangle TECAN Spark data. See flopr on github for more info!
#' @param data_csv needs to be the raw TECAN data in CSV form!
#' @param layout_csv needs to be filled in according to the example layout.
#'    you can add as many columns as you want BUT THE LAST COLUMN MUST BE named 'well' and have the position for each well such as 'A1'

plate1 <- drop_na(flopr::spark_parse(data_csv = "../Data/SF7_Probe/15Dec24_highOD_SF7.csv",
                                     layout_csv = "../Data/SF7_Probe/15Dec24_highOD_SF7_Metadata.csv",
                                     timeseries = TRUE))

#Change time if necessary! I am changing from seconds to hours!
plate1 %<>%
  mutate(across(time, ~.x / 3600)) %>%
  modify_if(is.character, as.factor) %>% 
  modify_at(c("Replicate"), as.factor)



readydata <- plate1 %>%
 # full_join(plate2) %>%
  #Wrangling data to make plots more easy to read.
  mutate(across(Plasmid, str_replace, 'pLW003', 'ev')) %>%
  mutate(across(Plasmid, str_replace, 'pLW002', '+SIR')) %>%
  mutate(Variant = Plasmid)
  

rm(plate1); #rm(plate2)
```

## Max and Min OD Graphs

### Quick visualization of data

```{r fig.height = 6, fig.width = 8}
#example of plotting raw data! I am only going to plot positive controls for this example!

readydata %>%
  ggplot(aes(x = time, y = SF7, color = Condition)) +
  geom_point(size = 2) + 
  labs(title = "raw data", x = "Hours", y = "SF7 RFU") +
  Theme

```

```{r}
readydata %>%
  group_by(Plasmid,Condition,time) %>%
  summarise(mean_FL = mean(SF7),
            SD_FL = sd(SF7)) -> data_summary
```

```{r, fig.height= 6, fig.width= 9}
data_summary %>%
  ggplot(aes(x = time, y = mean_FL, color = Condition)) +
  geom_point(size = 3) +
  geom_line(line_width = 3) +
  geom_errorbar(aes(ymin=mean_FL-SD_FL, ymax=mean_FL+SD_FL), width=.2,
                 position=position_dodge(0.05)) +
  labs(x = "Time (Hr)", y = "SF7 RFU") +
  facet_wrap(~Plasmid) +
  Theme
```


```{r, fig.height= 10, fig.width= 12}
readydata %>%
  ggplot(aes(x = time, y = SF7, color = Replicate)) +
  geom_point(size = 3) +
  stat_smooth(method = "lm", se = FALSE, color = "black") +
  ggpubr::stat_regline_equation(label.y.npc = .1, label.x.npc = .1, color = "black", size = 5) + 
  labs(x = "Time (Hr)", y = "SF7 RFU") +
  facet_wrap(Plasmid~Condition, ncol = 4) +
  Theme
```
Looks like there is no pattern between the different replicates which is good! There does to be a consistent bias in the data starting at approximately 2 hours. I am going to filter the data for that time!

```{r fig.height = 10, fig.width= 10}

p <- readydata %>%
  filter(time <= 1) %>%
  ggplot(aes(x = time, y = SF7)) +
  geom_point(size = 1.25, color = "#f0649a") +
  stat_smooth(method = "lm", se = FALSE, color = "black") +
  ggpubr::stat_regline_equation(label.y.npc = .1, label.x.npc = .1, color = "black", size = 5) + 
  labs(x = "Time (Hr)", y = "SF7 RFU", subtitle = "1hr") +
  facet_wrap(Plasmid~Condition, ncol = 4) +
  Theme

q <- readydata %>%
  filter(time <= .5) %>%
  ggplot(aes(x = time, y = SF7)) +
  geom_point(size = 1.25, color = "#f0649a") +
  stat_smooth(method = "lm", se = FALSE, color = "black") +
  ggpubr::stat_regline_equation(label.y.npc = .1, label.x.npc = .1, color = "black", size = 5) + 
  labs(x = "Time (Hr)", y = "SF7 RFU", subtitle = "30 min") +
  facet_wrap(Plasmid~Condition, ncol = 4) +
  Theme

p / q

```
This looks good and the slopes where the SIR is present seem to all be approximately linear, especially as the optical density grows! If I look at this data for only the first hour there is a large difference beteween the different slopes (11 data points). If I look at this at the thirty minute time scale, the slopes seem to become relatively indistinguishable between conditions. However, it's hard to tell if this is due to the conditions or because there is simply not enough data (only 7 points). I should retry this experiment for under and hour and take much for frequent time points. 

#Fitting individual rates for boxplots

The previous graphs were quick and dirty ways to quantify if the slopes were different between groups. Each replicate is a biological replicate and therefore requires it's own fitted line. The slope parameter can then be pulled out and placed into a boxplot! This could also be done by subtracting the 'ev' from the +SIR to get a difference between replicates! Let's calculate the slopes for all lines first and then we can subtract them and see how that changes!

```{r}
readydata %>%
  group_by(Strain, Plasmid, Condition, Replicate, Variant) %>%
  select(!c(Media,well, row, column)) %>%
  filter(time > .1 & time < 1) %>%
  nest(time, SF7) -> nested_data
```

```{r}
nested_data %<>%
  #fitting linear model
  mutate(model = map(data, function(df) lm(SF7 ~ time, data = df))) %>%
  #Create summary of coefficients
  mutate(summary = map(model, tidy)) %>%
  #Get residuals for model that can be plotted on top in a ggplot object
  mutate(augment = map(model, augment)) %>%
  #Using Glance to get an idea of model fit 
  mutate(model.fit = map(model, glance))
```

```{r fig.height = 6, fig.width = 9}
nested_data %>%
  unnest(augment) %>%
  ggplot(aes(x = time, y = SF7)) + 
  geom_point() + 
  #It's okay to change the x and y axis as long as you are in the same dataframe in this case. You are also changing the color of the global mapping and therefore you must include in the local aesthetic mapping to overwrite it!
  geom_line(aes(x = time, y = .fitted, color = Replicate), size = 1.25) +
  labs(y = "SF7 RFU", x = "Timepoint (Hr)") +
  facet_wrap(Plasmid~Condition, ncol = 4) +
Theme
```

## Converting Data to Boxplots

```{r}
nested_data %>%
  unnest(summary) %>%
  pivot_wider(names_from = term, values_from = c(estimate, std.error, statistic,p.value)) %>%
  group_by(Condition) %>%
  ggplot(aes(x = Condition, y = estimate_time, fill = Plasmid)) +
  geom_boxplot() + 
  geom_point(width = .4, position = position_dodge(width = .75), size = 1) +
  scale_y_continuous(limits = c(0,4000)) +
  labs(x = "Cell Concentration", y = "Slope coefficient") + 
  Theme
  
```
Wow these look great! I don't think I need to subtract the +SIR from the EV because I think there is such a dramatic difference at higher concentrations! I think this is the easiest way to show the data! Yay these look great! 
